---
title: "Step 4 - LDA output for labellers"
format: 
  html:
    number-sections: true
    anchor-sections: true
editor: visual
---

## LDA output for labellers

This template reads in the output from the LDA topic models and outputs materials needed for labelling:

-   Excel file per question/model with each response-topic assignment, along with gamma score and original text.

-   PDF file per question/model with top ten words and text from example responses for each topic.

Read in LDA output data and responses

```{r}
# Read the config file
config <- yaml::read_yaml(
  file.path("input", "config.yml")
)

general_params <- config$general_params

# Read in all responses
responses_all <- read_csv(
  paste0(
    general_params$output_dir,
    "responses_all_",
    str_replace_all(Sys.Date(), "-", "_"),
    ".csv"
  )
)
```

Create excel files

```{r}
lda_object <- readRDS("output/lda_object.rds")
lda_question_names <- lda_object %>% names()

for (question in lda_question_names) {
  
  # Get the gamma table for the current question
  table_gamma <- lda_object[[question]]$gamma
  
  # Join responses_all with the current table_gamma
  responses_and_gamma <- responses_all %>%
    select(response_id, question) %>% 
    mutate(response_id = as.character(response_id)) %>%
    left_join(table_gamma, by = c("response_id" = "document")) %>%
    mutate(question = question) %>% 
    rename(response_text = 2) %>% 
    select(
      response_id,
      question,
      response_text,
      topic,
      gamma
    )
  
  # Write to file
  filename <- paste0("output/output_for_labellers/topic_gamma_q", i, ".csv")
  write.csv(responses_and_gamma, file = filename, row.names = FALSE)
  
  # Progress message
  cat("Saved file:", filename, "\n")
}
```

Create PDF files by looping over \[output_for_labellers.rmd\](https://github.com/DataS-DHSC/consultation_example/blob/main/R/output_for_labellers.rmd)

```{r}
question_text_list <- names(lda_object)

for (i in seq_along(question_text_list)) {
  rmarkdown::render(
    "R/output_for_labellers.Rmd",
    output_file = paste0("../output/output_for_labellers/question_", i ,".pdf")
  )
}
```
